---
title: "Autumn Quarterly Project 2016"
output: html_document
---

The graphs below represent statistics relating to the [OpenStreetMap GB Autumn Quarterly Project 2016](https://wiki.openstreetmap.org/wiki/UK_2016_Q4_Project:_Food_Hygiene_Ratings), which involves using Food Hygiene Rating Scheme data to enrich the map. The data was gathered using the [FHRS/OSM comparison tool](http://gregrs.dev.openstreetmap.org/fhrs/) and relates to the time period between **18th October 2016** and **25th October 2016**. Please note that because of the OSM download process, there is a delay involved in new OSM data being processed by the comparison tool.

```{r load, include=T, echo=F}
old <- read.csv("~/Desktop/fhrs-stats/stats-2016-10-18.csv", row.names=2)
new <- read.csv("~/Desktop/fhrs-stats/stats-2016-10-25.csv", row.names=2)

# check length of old vs new CSVs to make sure that districts match up
# N.B. Haringey not present in CSVs before 2016-10-14
# To hide the output of this chunk in the HTML page, set include=F above
print("Number of districts matched between old and new CSVs?")
print((new$district_id - old$district_id)[length(new$district_id)] == 0)

# calculate percentage of matched FHRS establishments with an invalid/missing postcode
new$FHRS_pcode_err_pc <- (new$matched_postcode_error / new$total_FHRS) * 100
```

```{r prep-colours, echo=F}
blue <- "#c0c0ff"
green <- "#c0ffc0"
pink <- "#ffc0c0"

```

```{r prep-top10s, echo=F}
# difference in old vs new number and %age of matches
diffs <- data.frame(matched = new$matched - old$matched,
                    FHRS_matched_pc = new$FHRS_matched_pc - old$FHRS_matched_pc,
                    row.names=row.names(new))

# top 10s
top.matches.pc <- new[order(new$FHRS_matched_pc + new$FHRS_pcode_err_pc, decreasing=T),,
                      drop=F][1:10,,drop=F]
# we will use this matrix to create the stacked barplot
top.matches.pc.mtx <- t(as.matrix(top.matches.pc[,c("FHRS_matched_pc","FHRS_pcode_err_pc")]))
dimnames(top.matches.pc.mtx) = list(c("Valid postcode", "Mismatched postcode"),
                                    row.names(top.matches.pc))
top.pcode.pc <- new[order(new$OSM_matched_or_postcode_pc,
                          decreasing=T),,drop=F][1:10,,drop=F]
top.newmatch <- diffs[order(diffs$matched, decreasing=T),,drop=F][1:10,,drop=F]
top.newmatch.pc <- diffs[order(diffs$FHRS_matched_pc, decreasing=T),,drop=F][1:10,,drop=F]

# top mismatches and postcode errors
top.mismatch <- new[order(new$mismatch, decreasing=T),,drop=F][1:10,,drop=F]
top.pcode.error <- new[order(new$matched_postcode_error,
                             decreasing=T),,drop=F][1:10,,drop=F]
```

```{r prep-zero-matches, echo=F}
# how many districts have zero matches (old vs new)
no.match.districts = nrow(subset(new, matched == 0))
new.match.districts = nrow(subset(old, matched == 0)) - nrow(subset(new, matched == 0))
existing.match.districts = nrow(subset(old, matched > 0))
zero.matches.pie <- data.frame(districts = c(no.match.districts,
                                        new.match.districts,
                                        existing.match.districts),
                          row.names = c("zero matches",
                                        "now has >= 1 match",
                                        "already had >= 1 match"))
```

```{r prep-match-progress, echo=F}
# matching progress
new.match.pc <- (sum(new$matched) / sum(new$total_FHRS)) * 100
new.pcode.mismatch.pc <- (sum(new$matched_postcode_error) / sum(new$total_FHRS)) * 100
old.match.pc <- (sum(old$matched) / sum(new$total_FHRS)) * 100
old.pcode.mismatch.pc <- (sum(old$matched_postcode_error) / sum(new$total_FHRS)) * 100
progress.match.pc = new.match.pc - old.match.pc
progress.pcode.mismatch.pc = new.pcode.mismatch.pc - old.pcode.mismatch.pc

progress.numbers <- matrix(c(old.match.pc, old.pcode.mismatch.pc,
                             progress.match.pc, progress.pcode.mismatch.pc), nrow=2, ncol=2,
                             dimnames = list(c("Valid postcode", "Mismatched postcode"),
                                             c("Already matched", "Newly matched")))
```

<hr>

## Top 10 districts for completeness and growth

```{r plot-top10-matches, echo=F}
par(mai=c(1.1,2.5,0.75,0.75))
barplot(top.matches.pc.mtx[,10:1],
        horiz=T, las=1, col = c(blue,pink),
        main="Top 10: % of FHRS establishments matched",
        xlab="% of FHRS establishments matched",
        legend.text=T, args.legend = list(x=90, y=3.25, bty="n", cex=0.85))
```

```{r plot-top10-postcodes, echo=F}
par(mai=c(1.1,1.5,0.75,0.75))
barplot(rev(top.pcode.pc$OSM_matched_or_postcode_pc),
        names.arg = rev(row.names(top.pcode.pc)),
        horiz=T, las=1, col = blue,
        main="Top 10: % of relevant nodes/ways with a postcode",
        xlab="% of relevant nodes/ways with a non-mismatched postcode")
```

```{r plot-top10-new-matches, echo=F}
par(mai=c(1.1,2,0.75,0.75))
barplot(rev(top.newmatch$matched), names.arg = rev(row.names(top.newmatch)),
        horiz=T, las=1, col = green,
        main="Top 10: number of new matches",
        xlab="No. of newly matched FHRS establishments with valid postcodes")
```

```{r plot-top10-percentage-growth, echo=F}
par(mai=c(1.1,2,0.75,0.75))
barplot(rev(top.newmatch.pc$FHRS_matched_pc), names.arg = rev(row.names(top.newmatch.pc)),
        horiz=T, las=1, col = green,
        main="Top 10: growth in % of matches",
        xlab="Increase in % of matched FHRS establishments with valid postcodes")
```

## District completeness histograms

```{r plot-hist-matches, echo=F}
# histogram of district completeness %s
hist(new$FHRS_matched_pc + new$FHRS_pcode_err_pc, breaks=seq(0,100,5),
     main = "% of FHRS matches vs. number of districts",
     xlab = "% of FHRS establishments matched with or without postcode",
     ylab = "Number of districts",
     col = blue)
```

```{r plot-hist-postcode, echo=F}
hist(new$OSM_matched_or_postcode_pc, breaks=seq(0,100,5),
     main = "% of relevant nodes/ways with postcode vs. number of districts",
     xlab = "% of relevant OSM nodes/ways with postcode",
     ylab = "Number of districts",
     col = blue)
```

## Progress in matching FHRS establishments

```{r plot-match-progress, echo=F}
# bar plot of matching progress
barplot(progress.numbers,
        ylab = "% of FHRS establishments matched",
        col = c("#d0d0ff","#ffd0d0"),
        main = "Overall % of FHRS establishments matched",
        legend.text=T, args.legend = list(x=2.25, y=1.7, bty="n", cex=0.9))
```

```{r plot-zero-matches, echo=F}
# pie chart of districts with at least one match
par(mai=c(0.75,0.75,0.75,0.75))
lbls = paste(row.names(zero.matches.pie), " (", zero.matches.pie$districts, ")", sep="")
pie(zero.matches.pie$districts,
    labels = lbls,
    col = c("#ffd0d0","#d0ffd0","#d0d0ff"),
    main = "Districts with at least one match")
```

## Districts with the most errors

```{r plot-mismatches, echo=F}
par(mai=c(1.1,2.75,0.75,0.75))
barplot(rev(top.mismatch$mismatch), names.arg = rev(row.names(top.mismatch)),
        horiz=T, las=1, col = pink,
        main="Districts with the most FHRS ID mismatches",
        xlab="OSM nodes/ways with an unrecognised fhrs:id")
```

```{r plot-postcode-errors, echo=F}
par(mai=c(1.1,2.75,0.75,0.75))
barplot(rev(top.pcode.error$matched_postcode_error),
        names.arg = rev(row.names(top.pcode.error)),
        horiz=T, las=1, col = pink,
        main="Districts with the most missing/mismatched postcodes",
        xlab="Matched OSM nodes/ways with missing/mismatched postcode")
```

<hr>

*Created by Gregrs using [R](https://www.r-project.org) and [RStudio](https://www.rstudio.com). Code freely available on [GitHub](https://github.com/gregrs-uk/fhrs-osm-stats).*